An iterated Grothendieck construction of semi-cubical types

### 2022-02-23 paper/paper.tex
paper: major cleanup; removed derived notations
cohSheet is missing!

% Head and tail
\newcommand{\hd}{\ensuremath{\mathsf{hd}}}
\newcommand{\tl}{\ensuremath{\mathsf{tl}}}

% Universes, obj, mor
\newcommand{\sort}[1]{\ensuremath{\mathsf{U}_{#1}}}

@@

% Box, layer, and cube
\newcommand{\mycubset}[1]{\ensuremath{\mathsf{cubset}_{#1}}}
\newcommand{\mybox}[2]{\ensuremath{\mathsf{box}_{#1}^{#2}}}
\newcommand{\mylayer}[2]{\ensuremath{\mathsf{layer}_{#1}^{#2}}}
\newcommand{\mycube}[2]{\ensuremath{\mathsf{cube}_{#1}^{#2}}}

% Subbox, sublayer, and subcube
\newcommand{\downbox}[2]{\ensuremath{\mathsf{subbox}_{#1}^{#2}}}
\newcommand{\downlayer}[2]{\ensuremath{\mathsf{sublayer}_{#1}^{#2}}}
\newcommand{\downcube}[2]{\ensuremath{\mathsf{subcube}_{#1}^{#2}}}
\newcommand{\myfullbox}[2]{\ensuremath{\mathsf{fullbox}_{#1}^{#2}}}

@@

% Partials and fulls related to box, cube, and tube
\newcommand{\border}[1]{\ensuremath{\mathsf{border}_{#1}}}
\newcommand{\partialcubset}[2]{\ensuremath{\mathsf{cubset}_{#1}^{<#2}}}
\newcommand{\mycubsetcomp}[2]{\ensuremath{\mathsf{cubset}_{#1}^{=#2}}}
\newcommand{\fulldownbox}[2]{\ensuremath{\mathsf{subfullbox}_{#1}^{#2}}}
\newcommand{\fulldowncube}[2]{\ensuremath{\mathsf{subfullcube}_{#1}^{#2}}}
\newcommand{\propercube}[1]{\ensuremath{\mathsf{properfiller}_{#1}}}
\newcommand{\mygroundedcube}[1]{\ensuremath{\mathsf{groundedcube}_{#1}}}
\newcommand{\tube}[1]{\ensuremath{\mathsf{tube}_{#1}}}
\newcommand{\mycubsetfrom}[2]{\ensuremath{\mathsf{cubset}_{#1}^{\geq#2}}}
\newcommand{\myfullcube}[1]{\ensuremath{\mathsf{fullcube}_{#1}}}

@@

@@

\subsection{Truncated semi-cubical sets}
The definition is dispatched over tables \ref{tab:barecubicalsetstructurecore}, \ref{tab:barecubicalsetstructure}, \ref{tab:barecubicalsetfaces} and \ref{tab:barecubicalsetcoherences}. It describes the structure of the underlying higher-dimensional relations on which cubes and boxes are built, together with the definition of homogeneous $n$-boxes, homogeneous $n$-cubes, together with face operations on boxes and cubes, together with commutation properties of the face operations. All are mutually defined as types of the target language ETT. Note that such relational structure, cubes and boxes are relative to a universe. Note the presence of a coherence condition $\cohbox{l}{}$ to ensuring that both sides of the equality in $\downlayer{l}{}$ and $\downcube{l}{}$ are in the same type. The proof of $\cohbox{l}{}$ itself requires an higher-dimensional coherence condition which we obtain by working here in ETT where all proofs of an equality are identified (principle of Unicity of Identity Proofs). Note that if the proofs of the same equality were not equated, there would be a need for arbitrary many higher-dimensional coherences (see e.g.~\cite{Herbelin15} for a discussion on the de facto need for recursive higher-dimensional coherence conditions in formulating higher-dimensional structures in type theory). Note also that for a given $n$, the coherence conditions evaluate to a reflexivity proof, so that the construction evaluates to an effective sequence of types of iterated relations not mentioning $\downbox{l}{}$ nor $\cohbox{l}{}$ anymore.

@@

\begin{eqntable}{Definition of a truncated cubical set (main part of the higher-dimensional relation structure)\label{tab:barecubicalsetstructurecore}}
  \mc{Type of truncated cubical sets}
  \eqnline{\partialcubset{l}{n}}{}{:}{\sort{l+1}}
  \eqnline{\partialcubset{l}{0}}{}{\defeq}{\unittype}
  \eqnline{\partialcubset{l}{n'+1}}{}{\defeq}{\ensuremath{\Sigma}D:\partialcubset{l}{n'}.\,\mycubsetcomp{l}{n'}(D)}

@@

  \mc{Structure carried at each dimension}
  \eqnline{\mycubsetcomp{l}{n}}{(D:\partialcubset{l}{n})}{:}{\sort{l}}
  \eqnline{\mycubsetcomp{l}{n}}{D}{\defeq}{\myfullbox{l}{n}(D)\imp \sort{l}}
\end{eqntable}

\begin{eqntable}{Definition of a truncated cubical set (boxes)\label{tab:barecubicalsetstructure}}
  \mc{Full homogeneous $n$-boxes}
  \eqnline{\myfullbox{l}{n}}{(D:\partialcubset{l}{n})}{:}{\sort{l}}
  \eqnline{\myfullbox{l}{n}}{D}{\defeq}{\mybox{l}{n,n}(D)}

@@

  \mc{Homogeneous partial $n$-boxes and homogeneous partial $n$-cubes}
  \eqnline{\mybox{l}{n,p,[p \leq n]}}{(D:\partialcubset{l}{n})}{:}{\sort{l}}
  \eqnline{\mybox{l}{n,0}}{D}{\defeq}{\unittype}
  \eqnline{\mybox{l}{n,p'+1}}{D}{\defeq}{\Sigma d:\mybox{l}{n,p'}(D).\,\mylayer{l}{n,p'}(D)(d)}

@@

  \eqnline{\mylayer{l}{n,p,[p < n]}}{\makecell{(D:\partialcubset{l}{n}) \\ (d:\mybox{l}{n,p}(D))}}{:}{\sort{l}}

  \eqnline{\mylayer{l}{n,p}}{D~d}{\defeq}{\makecell{\mycube{l}{n-1,p}(\hd(D))(\tl(D))(\downbox{l,L,p}{n,p}(D)(d)) \\ \times\;\mycube{l}{n-1,p}(\hd(D))(\tl(D))(\downbox{l,R,p}{n,p}(D)(d))}}

@@

  \eqnline{\mycube{l}{n,p,[p \leq n]}}{\makecell{(D:\partialcubset{l}{n}) \\(E:\mycubsetcomp{l}{n}(D)) \\ (d:\mybox{l}{n,p}(D))}}{:}{\sort{l}}

  \eqnline{\mycube{l}{n,p,[p = n]}}{D~E~d}{\defeq}{E(d)}

  \eqnline{\mycube{l}{n,p,[p < n]}}{D~E~d}{\defeq}{\Sigma b:\mylayer{l}{n,p}(D)(d).\,\mycube{l}{n,p+1}(D)(E)(d,b)}
\end{eqntable}

@@

  \eqnline{\downbox{l,\epsilon,q}{n,p,[p \leq q \leq n - 1]}}{\makecell{(D:\partialcubset{l}{n}) \\ (d:\mybox{l}{n,p}(D))}}{:}{\mybox{l}{n-1,p}(\hd(D))}

  \eqnline{\downbox{l,\epsilon,q}{n,p'+1}}{D~(d,b)}{\defeq}{(\downbox{l,\epsilon,q}{n,p'}(D)(d),\downlayer{l,\epsilon,q-1}{n,p'}(D)(d)(b))}

@@

  \eqnline{\downlayer{l,\epsilon,q}{n,p,[p \leq q \leq n - 2]}}{\makecell{(D:\partialcubset{l}{n}) \\ (d:\mybox{l}{n,p}(D)) \\ (b:\mylayer{l}{n,p}(D)(d))}}{:}{\mylayer{l}{n-1,p}(\hd(D))(\downbox{l,\epsilon,q+1}{n,p}(D)(d))}

  \eqnline{\downlayer{l,\epsilon,q}{n,p}}{D~d~c}{\defeq}{\makecell{(\overright{\cohbox{l,\epsilon,L,q,p}{n,p}(D)(d)}(\downcube{l,\epsilon,q}{n-1,p}(\hd(D))(\tl(D))(\downbox{l,L,p}{n,p}(D)(d))(c_L)), \\ \;\overright{\cohbox{l,\epsilon,R,q,p}{n,p}(D)(d)}(\downcube{l,\epsilon,q}{n-1,p}(\hd(D))(\tl(D))(\downbox{l,R,p}{n,p}(D)(d))(c_R)))}}

@@

  \eqnline{\downcube{l,\epsilon,q}{n,p,[p \leq q \leq n - 1]}}{\makecell{(D:\partialcubset{l}{n}) \\ (E:\mycubsetcomp{l}{n}(D)) \\(d:\mybox{l}{n,p}(D)) \\ (b:\mycube{l}{n,p}(D)(E)(d))}}{:}{\mycube{l}{n-1,p}(\hd(D))(\tl(D))(\downbox{l,\epsilon,q+1}{n,p}(D)(d))}

  \eqnline{\downcube{l,\epsilon,q}{n,p,[p=q]}}{D~E~d~(b,\_)}{\defeq}{b_\epsilon}

  \eqnline{\downcube{l,\epsilon,q}{n,p,[p<q]}}{D~E~d~(b,c)}{\defeq}{(\downlayer{l,\epsilon,q}{n,p}(D)(d)(b),\downcube{l,\epsilon,q}{n,p+1}(D)(E)(d,b)(c))}
\end{eqntable}

@@

  \eqnline{\cohbox{l,\epsilon,\omega,q,r}{n,p,[p \leq r \leq q \leq n - 2]}}{\makecell{(D:\partialcubset{l}{n}) \\ (d:\mybox{l}{n,p}(D))}}{:}{\makecell{\downbox{l,\epsilon,q}{n-1,p}(\hd(D))(\downbox{l,\omega,r}{n,p}(D)(d)) \\ \eqett \downbox{l,\omega,r}{n-1,p}(\hd(D))(\downbox{l,\epsilon,q+1}{n,p}(D)(d))}}

@@

  \eqnline{\cohbox{l,\epsilon,\omega,q,r}{n,p'+1}}{D~(d,b)}{\defeq}{(\cohbox{l,\epsilon,\omega,q,r}{n,p'}(D)(d),\cohlayer{l,\epsilon,\omega,q,r}{n,p'}(D)(d)(b))}

@@

  \eqnline{\cohlayer{l,\epsilon,\omega,q,r}{n,p,[p < r \leq q \leq n - 2]}}{\makecell{(D:\partialcubset{l}{n}) \\ (d:\mybox{l}{n,p}(D)) \\(b:\mylayer{l}{n,p}(D)(d))}}{:}{\makecell{\overright{\cohbox{l,\epsilon,\omega,q+1,r+1}{n,p}(D)(d)}(\downlayer{l,\epsilon,q}{n-1,p}(\hd(D))(\downbox{l,\omega,r}{n,p}(D)(d))(\downlayer{l,\omega,r}{n,p}(D)(d)(b))) \\ \eqett \downlayer{l,\omega,r}{n-1,p}(\hd(D))(\downbox{l,\epsilon,q+1}{n,p}(D)(d))(\downlayer{l,\epsilon,q+1}{n,p}(D)(d)(b))}}

  \eqnline{\cohlayer{l,\epsilon,\omega,q,r}{n,p}}{D~d~c}{\defeq}{\makecell{(\cohcube{l,\epsilon,\omega,q-1,r-1}{n-1,p}(\hd(D))(\tl(D))(\downbox{l,L,p}{n,p}(D)(d))(c_L), \\ \;\cohcube{l,\epsilon,\omega,q-1,r-1}{n-1,p}(\hd(D))(\tl(D))(\downbox{l,R,p}{n,p}(D)(d))(c_R))}}

@@

  \eqnline{\cohcube{l,\epsilon,\omega,q,r}{n,p,[p \leq r \leq q \leq n - 2]}}{\makecell{(D:\partialcubset{l}{n}) \\ (E:\mycubsetcomp{l}{n}(D)) \\ (d:\mybox{l}{n,p}(D)) \\ (b:\mycube{l}{n,p}(D)(E)(d))}}{:}{\makecell{\overright{\cohbox{l,\epsilon,\omega,q+1,r+1}{n,p}(D)(d)}(\downcube{l,\epsilon,q}{n-1,p}(\hd(D))(\tl(D))(\downbox{l,\omega,r}{n,p}(D)(d))(\downcube{l,\omega,r}{n,p}(D)(E)(d)(b))) \\ \eqett \downcube{l,\omega,r}{n-1,p}(\hd(D))(\tl(D))(\downbox{l,\epsilon,q+1}{n,p}(D)(d))(\downcube{l,\epsilon,q+1}{n,p}(D)(E)(d)(b))}}

  \eqnline{\cohcube{l,\epsilon,\omega,q,r}{n,p,[p=r]}}{D~E~d~(b,\_)}{\defeq}{\reflett(\downcube{l,\epsilon,q-1}{n-1,p}(\hd(D))(\tl(D))(\downbox{l,\omega,p}{n,p}(D)(d))(b_{\epsilon}))}

  \eqnline{\cohcube{l,\epsilon,\omega,q,r}{n,p,[p<r]}}{D~E~d~(b,c)}{\defeq}{(\cohlayer{l,\epsilon,\omega,q,r}{n,p}(D)(d)(b),\;\cohcube{l,\epsilon,\omega,q,r}{n,p+1}(D)(E)(d,b)(c))}
\end{eqntable}

\subsection{Derived notions\label{sec:derived-notions}}
On top of the basic definitions on tables \ref{tab:barecubicalsetstructure}, \ref{tab:barecubicalsetfaces} and \ref{tab:barecubicalsetcoherences}, we can define on tables \ref{tab:macros} and \ref{tab:macroscont} a few other concepts at a more standard level of abstraction. In particular, a homogeneous $n$-box in the above sense is defined to be an object of type $\myfullbox{l}{n}(D_{n})$ for $D_{n} \defeq (A, =_A, \ldots, =^n_A)$ the initial segment of $n$-th first iterated equalities over $A$ (by convention, a homogeneous $0$-box shall be the canonical singleton type). A $n$-tube is of type $\mybox{l}{n+1,n}(D_{n+1})$. An homogeneous (proper) $n$-cube in some $n$-box $d$ is of type $\mycube{l}{n+1,n}(\hd(D_{n+1}))(\tl(D_{n+1}))(d)$ and a full $n$-cube is of type $\mycube{l}{n+1,0}(\hd(D_{n+1}))(\tl(D_{n+1})))(\star)$.

Note that the components of an $n$-box or $n$-tube are tuples associated to the left. For cubes, the components can be associated to the left, and we call this a full $n$-cube or to the right, which we call grounded $n$-cube, and which corresponds to ``complete'' partial cubes, i.e. to partial cubes over the empty box. Table \ref{tab:macros} show how to compute the $n$-box surrounding a full $n$-cube over a prefix $D_{n+1}$ of iterated equalities (see $\border{}$).

Table \ref{tab:macros} also shows how to compute the faces of a box or of a cube.

\begin{eqntable}{Definition of a bare cubical set (coinductive structure)\label{tab:barecubicalset}}
  \mc{Full (non-truncated) cubical sets with degeneracies}
  \eqnline{\mycubset{l}}{}{:}{\sort{l+1}}
  \eqnline{\mycubset{l}}{}{\defeq}{\mycubsetfrom{l}{0}(\star)}
  \\
  \eqnline{\mycubsetfrom{l}{n}}{(D:\partialcubset{l}{n})}{:}{\sort{l+1}}
  \eqnline{\mycubsetfrom{l}{n}}{D}{\defeq}{\Sigma R:\mycubsetcomp{l}{n}(D).\,\mycubsetfrom{l}{n+1}(D,R)}
\end{eqntable}

\begin{eqntable}{Standard derived notions\label{tab:macros}}
  \mc{Type of full $n$-cubes}
  \eqnline{\myfullcube{l}^{n}}{(D:\partialcubset{l}{n+1})}{:}{\sort{l}}

  \eqnline{\myfullcube{l}^{n}}{(D,E)}{\defeq}{\Sigma d:\myfullbox{l}{n}(D).\, E.\mathsf{rel}(d)}

  \\ \mc{Type of $n$-tubes}

  \eqnline{\tube{l}^{n}}{(D:\partialcubset{l}{n+1})}{:}{\sort{l}}
  \eqnline{\tube{l}^{n}}{D}{\defeq}{\mybox{l}{n+1,n}(D)}

  \\ \mc{Type of grounded $n$-cubes}

  \eqnline{\mygroundedcube{l}^{n}}{(D:\partialcubset{l}{n+1})}{:}{\sort{l}}
  \eqnline{\mygroundedcube{l}^{n}}{(D,E)}{\defeq}{\mycube{l}{n,0}(D)(E)(\star)}

  \\ \mc{Type of fillers of some $n$-box}

  \eqnline{\propercube{l}^{n}}{(D:\partialcubset{l}{n+1})~(d:\myfullbox{l}{n}(\hd(D)))}{:}{\sort{l}}

  \eqnline{\propercube{l}^{n}}{(D,E)~d}{\defeq}{E.\mathsf{rel}(d)}

  \\ \mc{Border of a grounded $n$-cube}

  \eqnline{\border{l}^{n}}{(D:\partialcubset{l}{n+1})~(c:\mygroundedcube{l}^{n}(D))}{:}{\myfullbox{l}{n}(\hd(D))}

  \eqnline{\border{l}^{n}}{D~c}{\defeq}{\border{l}^{n,0}(D)(\star)(c)}

  \\ \mc{where the border of partial cubes is defined by}

  \eqnline{\border{l}^{n,p}}{(D:\partialcubset{l}{n+1})~(d:\mybox{l}{n,p}(\hd(D)))~(c:\mycube{l}{n,p}(\hd(D))(\tl(D))(d))}{:}{\myfullbox{l}{n}(\hd(D))}

  \eqnline{\border{l}^{n,p,[p=n]}}{D~d~c}{\defeq}{d}

  \eqnline{\border{l}^{n,p,[p<n]}}{D~d~(b,c)}{\defeq}{\border{l}^{n,p+1}(D)(d,b)(c)}
\end{eqntable}

\begin{eqntable}{Standard derived notions (continued)\label{tab:macroscont}}
  \mc{$q$-th projection of an $n$-box}

  \eqnline{\fulldownbox{l,\epsilon,q}{n,[q < n]}}{(D:\partialcubset{l}{n})~(d:\myfullbox{l}{n}(D))}{:}{\myfullbox{l}{n-1}(\hd(D))}

  \eqnline{\fulldownbox{l,\epsilon,q}{n}}{D~(d,\_)}{\defeq}{\downbox{l,\epsilon,q}{n,n-1}(D)(d)}

  \\ \mc{$q$-th face of an $n$-cube}

  \eqnline{\fulldowncube{l,\epsilon,q}{n,[q < n]}}{(D:\partialcubset{l}{n+1})~(c:\myfullcube{l}^{n}(D))}{:}{\myfullcube{l}^{n-1}(\hd(D))}

  \eqnline{\fulldowncube{l,\epsilon,q}{n}}{(D,E)~c}{\defeq}{\downcube{l,\epsilon,q}{n,0}(D)(E)(\star)(c)}

  \\ \mc{where the extension of $\downbox{l}{}$ to $p>q$ is}

  \eqnline{\downbox{l,\epsilon,q}{n,p,[q < p < n]}}{(D:\partialcubset{l}{n})~(d:\mybox{l}{n,p}(D))}{:}{\myfullbox{l}{n-1}(\hd(D))}

  \eqnline{\downbox{l,\epsilon,q}{n,q+1}}{D~(d,b)}{\defeq}{\border{l}^{n-1,q}(\hd(D))(\downbox{l,\epsilon,q}{n,q}(D)(d))(b_{\epsilon})}

  \eqnline{\downbox{l,\epsilon,q}{n,p,[q < p-1]}}{D~(d,b)}{\defeq}{\downbox{l,\epsilon,q}{n,p-1}(D)(d)}
\end{eqntable}

@@

    Definition le := le'.
    \end{minted}
    \end{listing}
  \item[(c)] A refinement of the inductive definition; a nested implication.
    \begin{listing}[H]
      \begin{minted}{coq}
      Definition le n m := forall p, p <= n -> p <= m.
      \end{minted}
    \end{listing}
\end{enumerate}

(a) was unworkable right from the start, and we tried approach (b) for a while, before running into unification problems. We then settled on approach (c), because it made compositions of inequalities almost trivial to prove. Compare the following snippets defining transitivity of $\leq$:

>>>>

% Universes
\newcommand{\sort}[1]{\ensuremath{\mathsf{U}_{#1}}}

@@

% cubset
\newcommand{\csp}[1]{\ensuremath{\mathsf{cubset}_{#1}}}
\newcommand{\cubeset}[2]{\ensuremath{\mathsf{cubset}_{#1}^{<#2}}}
\newcommand{\cubesetcomp}[2]{\ensuremath{\mathsf{cubset}_{#1}^{=#2}}}
\newcommand{\cubesetfrom}[2]{\ensuremath{\mathsf{cubset}_{#1}^{\geq#2}}}

% Box, layer, and cube
\newcommand{\mybox}[2]{\ensuremath{\mathsf{box}_{#1}^{#2}}}
\newcommand{\layer}[2]{\ensuremath{\mathsf{layer}_{#1}^{#2}}}
\newcommand{\cube}[2]{\ensuremath{\mathsf{cube}_{#1}^{#2}}}

% Subbox, sublayer, and subcube
\newcommand{\subbox}[2]{\ensuremath{\mathsf{subbox}_{#1}^{#2}}}
\newcommand{\sublayer}[2]{\ensuremath{\mathsf{sublayer}_{#1}^{#2}}}
\newcommand{\subcube}[2]{\ensuremath{\mathsf{subcube}_{#1}^{#2}}}
\newcommand{\fullbox}[2]{\ensuremath{\mathsf{fullbox}_{#1}^{#2}}}

@@

@@

% Some abbreviations
\newcommand{\D}{\ensuremath{(D)}}
\newcommand{\hdD}{\ensuremath{(D.1)}}
\newcommand{\tlD}{\ensuremath{(D.2)}}
\renewcommand{\d}{\ensuremath{(d)}}
\newcommand{\E}{\ensuremath{(E)}}
\renewcommand{\l}{\ensuremath{(l)}}
\renewcommand{\c}{\ensuremath{(c)}}

@@

\subsection{Truncated semi-cubical sets}
The definition is dispatched over tables \ref{tab:barecubicalsetstructurecore}, \ref{tab:barecubicalsetstructure}, \ref{tab:barecubicalsetfaces} and \ref{tab:barecubicalsetcoherences}. It describes the structure of the underlying higher-dimensional relations on which cubes and boxes are built, together with the definition of homogeneous $n$-boxes, homogeneous $n$-cubes, together with face operations on boxes and cubes, together with commutation properties of the face operations. All are mutually defined as types of the target language ETT. Note that such relational structure, cubes and boxes are relative to a universe. Note the presence of a coherence condition $\cohbox{l}{}$ to ensuring that both sides of the equality in $\sublayer{l}{}$ and $\subcube{l}{}$ are in the same type. The proof of $\cohbox{l}{}$ itself requires an higher-dimensional coherence condition which we obtain by working here in ETT where all proofs of an equality are identified (principle of Unicity of Identity Proofs). Note that if the proofs of the same equality were not equated, there would be a need for arbitrary many higher-dimensional coherences (see e.g.~\cite{Herbelin15} for a discussion on the de facto need for recursive higher-dimensional coherence conditions in formulating higher-dimensional structures in type theory). Note also that for a given $n$, the coherence conditions evaluate to a reflexivity proof, so that the construction evaluates to an effective sequence of types of iterated relations not mentioning $\subbox{l}{}$ nor $\cohbox{l}{}$ anymore.

@@

\begin{eqntable}{Definition of a truncated cubical set (main part of the higher-dimensional relation structure)\label{tab:barecubicalsetstructurecore}}
  \mc{Type of truncated cubical sets}
  \eqnline{\cubeset{l}{n}}{}{:}{\sort{l+1}}
  \eqnline{\cubeset{l}{0}}{}{\defeq}{\unittype}
  \eqnline{\cubeset{l}{n'+1}}{}{\defeq}{\ensuremath{\Sigma}D:\cubeset{l}{n'}.\,\cubesetcomp{l}{n'} \D}

@@

  \mc{Structure carried at each dimension}
  \eqnline{\cubesetcomp{l}{n}}{(D:\cubeset{l}{n})}{:}{\sort{l}}
  \eqnline{\cubesetcomp{l}{n}}{D}{\defeq}{\fullbox{l}{n}\D\imp \sort{l}}
\end{eqntable}

\begin{eqntable}{Definition of a truncated cubical set (boxes)\label{tab:barecubicalsetstructure}}
  \mc{Full homogeneous $n$-boxes}
  \eqnline{\fullbox{l}{n}}{(D:\cubeset{l}{n})}{:}{\sort{l}}
  \eqnline{\fullbox{l}{n}}{D}{\defeq}{\mybox{l}{n,n}\D}

@@

  \mc{Homogeneous partial $n$-boxes and homogeneous partial $n$-cubes}
  \eqnline{\mybox{l}{n,p,[p \leq n]}}{(D:\cubeset{l}{n})}{:}{\sort{l}}
  \eqnline{\mybox{l}{n,0}}{D}{\defeq}{\unittype}
  \eqnline{\mybox{l}{n,p'+1}}{D}{\defeq}{\Sigma d:\mybox{l}{n,p'}\D.\,\layer{l}{n,p'} \D \d}

@@

  \eqnline{\layer{l}{n,p,[p < n]}}{\makecell{(D:\cubeset{l}{n}) \\ (d:\mybox{l}{n,p} \D)}}{:}{\sort{l}}

  \eqnline{\layer{l}{n,p}}{D~d}{\defeq}{\makecell{\cube{l}{n-1,p}\hdD\tlD(\subbox{l,L,p}{n,p} \D \d) \\ \times\;\cube{l}{n-1,p}\hdD\tlD(\subbox{l,R,p}{n,p} \D \d)}}

@@

  \eqnline{\cube{l}{n,p,[p \leq n]}}{\makecell{(D:\cubeset{l}{n}) \\(E:\cubesetcomp{l}{n} \D) \\ (d:\mybox{l}{n,p} \D)}}{:}{\sort{l}}

  \eqnline{\cube{l}{n,p,[p = n]}}{D~E~d}{\defeq}{\E \d}

  \eqnline{\cube{l}{n,p,[p < n]}}{D~E~d}{\defeq}{\Sigma l:\layer{l}{n,p} \D \d.\,\cube{l}{n,p+1} \E(d,l)}
\end{eqntable}

@@

  \eqnline{\subbox{l,\epsilon,q}{n,p,[p \leq q \leq n - 1]}}{\makecell{(D:\cubeset{l}{n}) \\ (d:\mybox{l}{n,p} \D)}}{:}{\mybox{l}{n-1,p}\hdD}

  \eqnline{\subbox{l,\epsilon,q}{n,p'+1}}{D~(d,l)}{\defeq}{(\subbox{l,\epsilon,q}{n,p'} \D\d,\sublayer{l,\epsilon,q-1}{n,p'} \D \d\l)}

@@

  \eqnline{\sublayer{l,\epsilon,q}{n,p,[p \leq q \leq n - 2]}}{\makecell{(D:\cubeset{l}{n}) \\ (d:\mybox{l}{n,p} \D) \\ (b:\layer{l}{n,p} \D \d)}}{:}{\layer{l}{n-1,p}\hdD(\subbox{l,\epsilon,q+1}{n,p} \D \d)}

  \eqnline{\sublayer{l,\epsilon,q}{n,p}}{D~d~c}{\defeq}{\makecell{(\overright{\cohbox{l,\epsilon,L,q,p}{n,p} \D \d}(\subcube{l,\epsilon,q}{n-1,p}\hdD\tlD(\subbox{l,L,p}{n,p} \D \d)(c_L)), \\ \;\overright{\cohbox{l,\epsilon,R,q,p}{n,p} \D \d}(\subcube{l,\epsilon,q}{n-1,p}\hdD\tlD(\subbox{l,R,p}{n,p}\D\d)(c_R)))}}

@@

  \eqnline{\subcube{l,\epsilon,q}{n,p,[p \leq q \leq n - 1]}}{\makecell{(D:\cubeset{l}{n}) \\ (E:\cubesetcomp{l}{n} \D) \\(d:\mybox{l}{n,p} \D) \\ (b:\cube{l}{n,p} \D\E\d)}}{:}{\cube{l}{n-1,p}\hdD\tlD(\subbox{l,\epsilon,q+1}{n,p} \D \d)}

  \eqnline{\subcube{l,\epsilon,q}{n,p,[p=q]}}{D~E~d~(b,\_)}{\defeq}{b_\epsilon}

  \eqnline{\subcube{l,\epsilon,q}{n,p,[p<q]}}{D~E~d~(b,c)}{\defeq}{(\sublayer{l,\epsilon,q}{n,p} \D \d(b),\subcube{l,\epsilon,q}{n,p+1} \D\E(d,b)\c)}
\end{eqntable}

@@

  \eqnline{\cohbox{l,\epsilon,\omega,q,r}{n,p,[p \leq r \leq q \leq n - 2]}}{\makecell{(D:\cubeset{l}{n}) \\ (d:\mybox{l}{n,p} \D)}}{:}{\makecell{\subbox{l,\epsilon,q}{n-1,p}\hdD(\subbox{l,\omega,r}{n,p} \D \d) \\ \eqett \subbox{l,\omega,r}{n-1,p}\hdD(\subbox{l,\epsilon,q+1}{n,p} \D \d)}}

@@

  \eqnline{\cohbox{l,\epsilon,\omega,q,r}{n,p'+1}}{D~(d,c)}{\defeq}{(\cohbox{l,\epsilon,\omega,q,r}{n,p'} \D\d,\cohlayer{l,\epsilon,\omega,q,r}{n,p'} \D\d\c)}

@@

  \eqnline{\cohlayer{l,\epsilon,\omega,q,r}{n,p,[p < r \leq q \leq n - 2]}}{\makecell{(D:\cubeset{l}{n}) \\ (d:\mybox{l}{n,p} \D) \\(b:\layer{l}{n,p} \D \d)}}{:}{\makecell{\overright{\cohbox{l,\epsilon,\omega,q+1,r+1}{n,p} \D \d}(\sublayer{l,\epsilon,q}{n-1,p}\hdD(\subbox{l,\omega,r}{n,p} \D \d)(\sublayer{l,\omega,r}{n,p} \D \d(b))) \\ \eqett \sublayer{l,\omega,r}{n-1,p}\hdD(\subbox{l,\epsilon,q+1}{n,p} \D \d)(\sublayer{l,\epsilon,q+1}{n,p} \D \d(b))}}

  \eqnline{\cohlayer{l,\epsilon,\omega,q,r}{n,p}}{D~d~c}{\defeq}{\makecell{(\cohcube{l,\epsilon,\omega,q-1,r-1}{n-1,p}\hdD\tlD(\subbox{l,L,p}{n,p} \D \d)(c_L), \\ \;\cohcube{l,\epsilon,\omega,q-1,r-1}{n-1,p}\hdD\tlD(\subbox{l,R,p}{n,p} \D \d)(c_R))}}

@@

  \eqnline{\cohcube{l,\epsilon,\omega,q,r}{n,p,[p \leq r \leq q \leq n - 2]}}{\makecell{(D:\cubeset{l}{n}) \\ (E:\cubesetcomp{l}{n} \D) \\ (d:\mybox{l}{n,p} \D) \\ (b:\cube{l}{n,p} \D\E \d)}}{:}{\makecell{\overright{\cohbox{l,\epsilon,\omega,q+1,r+1}{n,p} \D \d}(\subcube{l,\epsilon,q}{n-1,p}\hdD\tlD(\subbox{l,\omega,r}{n,p} \D \d)(\subcube{l,\omega,r}{n,p} \D\E \d(b))) \\ \eqett \subcube{l,\omega,r}{n-1,p}\hdD\tlD(\subbox{l,\epsilon,q+1}{n,p} \D \d)(\subcube{l,\epsilon,q+1}{n,p} \D\E \d(b))}}

  \eqnline{\cohcube{l,\epsilon,\omega,q,r}{n,p,[p=r]}}{D~E~d~(b,\_)}{\defeq}{\reflett(\subcube{l,\epsilon,q-1}{n-1,p}\hdD\tlD(\subbox{l,\omega,p}{n,p} \D \d)(b_{\epsilon}))}

  \eqnline{\cohcube{l,\epsilon,\omega,q,r}{n,p,[p<r]}}{D~E~d~(b,c)}{\defeq}{(\cohlayer{l,\epsilon,\omega,q,r}{n,p} \D \d(b),\;\cohcube{l,\epsilon,\omega,q,r}{n,p+1} \D\E(d,b)\c)}
\end{eqntable}

\begin{eqntable}{Definition of a bare cubical set (coinductive structure)\label{tab:barecubicalset}}
  \mc{Full (non-truncated) cubical sets with degeneracies}
  \eqnline{\csp{l}}{}{:}{\sort{l+1}}
  \eqnline{\csp{l}}{}{\defeq}{\cubesetfrom{l}{0}(\star)}
  \\
  \eqnline{\cubesetfrom{l}{n}}{(D:\cubeset{l}{n})}{:}{\sort{l+1}}
  \eqnline{\cubesetfrom{l}{n}}{D}{\defeq}{\Sigma R:\cubesetcomp{l}{n} \D.\,\cubesetfrom{l}{n+1}(D,R)}
\end{eqntable}

@@

    Definition le := le'.
    \end{minted}
    \end{listing}
  \item[\c] A refinement of the inductive definition; a nested implication.
    \begin{listing}[H]
      \begin{minted}{coq}
      Definition le n m := forall p, p <= n -> p <= m.
      \end{minted}
    \end{listing}
\end{enumerate}

(a) was unworkable right from the start, and we tried approach (b) for a while, before running into unification problems. We then settled on approach \c, because it made compositions of inequalities almost trivial to prove. Compare the following snippets defining transitivity of $\leq$:
